A linearization procedure for constrained multibody systems

### 2012-06-22 PetersonGedeHubbard2012.tex
Added rolling disc figure.

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
%\usepackage[polutonikogreek]{babel}

@@

Currently, there are only a few approaches to linearizing constrained systems
which have been presented (citations here). A limitation of these approaches is
that they have been developed for equations of motion found using Lagrange¿s
method, which contain Lagrange multipliers. Kane¿s method (citation here)
allows for formulation of equations of motion for constrained systems in which
no multipliers are present. The dynamics of these systems can be completely
described by ordinary differential equations. This significantly different
formulation requires a different approach to linearization than what the
existing options provide.

The goal of the following paper is to concretely establish the first order
relationships (about an arbitrary equilibrium point) between a selection of
independent coordinates and speeds and the time-derivatives of all of the
coordinates and speeds (both independent and dependent). The formalism
presented here is generic enough to cover most examples of time-varying
constrained multibody systems with arbitrary external inputs and arbitrary
specified quantities.  This procedure was created for systems whose equations
of motion have been derived with Kane¿s Method. Although also applicable to
dynamic system equations formulated using other methods, it is restricted to
systems which can be completely described by a set of ODEs (i.e., no DAEs are
present).

We first present our linearization procedure. Next we will offer some
discussion of the procedure, along with an example. The derivation of the
procedure is provided in the appendix. Additionally, the electronic
supplementary material provides more examples (in more detail) and an
alternative ordering for the independent/dependent quantities.

@@

\section{Results}
\label{sec:results}
The linearized equations of motion are
\begin{align}
  \label{eq:state_space_constrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q}_i \\
        \delta \bm{u}_i
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n+m) \times s} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}
Table \ref{table:assumptions} provides the form in which the nonlinear
equations of motion and constraints must be arranged.  Equations
\ref{eq:quant_to_compute} provides the bulk of the terms in equation
\ref{eq:state_space_constrained}; these terms are Jacobian matrices of the
differential equations. Equations \ref{eq:C_0}-\ref{eq:C_2} provide the terms
$C_0$, $C_1$, and $C_2$ which embed the constraints into
\ref{eq:state_space_constrained}. Finally, the permutation matrices in
equations \ref{eq:Pq} and \ref{eq:Pu} provide the means to separate the
independent and dependent speeds.

@@

\section{Discussion}
\label{sec:discussion}
Equation \ref{eq:state_space_constrained} doesn't fit into the standard linear
system framework because it is non-square.  To obtain a square system (useful
in stability analysis and control system design) we define
\begin{align}
  \label{eq:A_prime}
    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
  \label{eq:B_prime}
    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
\end{align}
where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
independent states by defining
\begin{align}
  \label{eq:P_prime}
    P^\prime &\triangleq \begin{bmatrix}
        P_{qi} & \bm{O}_{n \times (o - m)} \\
        \bm{O}_{o \times (n - l)} & P_{ui}
    \end{bmatrix} \\
  \label{eq:A}
    A &\triangleq P^{\prime T} A^\prime \\
  \label{eq:B}
    B &\triangleq P^{\prime T} B^\prime
\end{align}
where $P^\prime \in \bm{R}^{(o - m + n - l) \times (o - m + n - l)}$.  Defining
$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
linear systems analyses may applied.  It is worth noting that the rows
of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
in the output or measurement equations of a linear state space model, as in the
case when the controlling or measuring dependent state(s).

@@

%f_c:
%[r*cos(q1) + q5]
%f_v:
%[                r*u1*cos(q2) + u3]
%[-r*(u0*cos(q2) + u2*sin(q2)) + u4]
%[                r*u1*sin(q2) + u5]
%f_a:
%[                                -r*u1*sin(q2)*q2' + r*cos(q2)*u1' + u3']
%[-r*(-u0*sin(q2)*q2' + u2*cos(q2)*q2' + sin(q2)*u2' + cos(q2)*u0') + u4']
%[                                 r*u1*cos(q2)*q2' + r*sin(q2)*u1' + u5']
%f_0:
%[                                                                                      -sin(q2)*cos(q1)*q0' + cos(q2)*q1']
%[                                                                                                       sin(q1)*q0' + q2']
%[                                                                                       sin(q2)*q1' + cos(q1)*cos(q2)*q0']
%[(sin(q0)*cos(q2) + sin(q1)*sin(q2)*cos(q0))*q4' + (-sin(q0)*sin(q1)*sin(q2) + cos(q0)*cos(q2))*q3' - sin(q2)*cos(q1)*q5']
%[                                                                -sin(q0)*cos(q1)*q3' + sin(q1)*q5' + cos(q0)*cos(q1)*q4']
%[ (sin(q0)*sin(q2) - sin(q1)*cos(q0)*cos(q2))*q4' + (sin(q0)*sin(q1)*cos(q2) + sin(q2)*cos(q0))*q3' + cos(q1)*cos(q2)*q5']
%f_1:
%[-u0]
%[-u1]
%[-u2]
%[-u3]
%[-u4]
%[-u5]
%f_2:
%[                  -I*u0' - m*r*cos(q2)*u4']
%[-J*u1' + m*r*sin(q2)*u5' + m*r*cos(q2)*u3']
%[                  -I*u2' - m*r*sin(q2)*u4']
%f_3:
%[-I*u1*u2 + J*u1*u2 + g*m*r*sin(q1)*cos(q2) - m*r*(-u0*u5 + u2*u3)*cos(q2)]
%[                m*r*(u0*u4 - u1*u3)*sin(q2) + m*r*(u1*u5 - u2*u4)*cos(q2)]
%[ I*u0*u1 - J*u0*u1 + g*m*r*sin(q1)*sin(q2) - m*r*(-u0*u5 + u2*u3)*sin(q2)]
\section{Derivation of linearization algorithm}
\label{sec:derivations}

This section outlines the derivation of the procedure for determining linear
equations of motion \ref{eq:state_space_constrained}. This equation can be used
for simulation, stability analysis, or it can be reformulated into the
traditional state space equations for a variety of tasks; these topics are
discussed in \ref{sec:discussion}.

>>>>

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{pgf}
%\usepackage[polutonikogreek]{babel}

@@

Currently, there are only a few approaches to linearizing constrained systems
which have been presented\cite{Kang2003}, \cite{Negrut2006}. A limitation of these approaches is
that they have been developed for equations of motion found using Lagrange's
method, which contain Lagrange multipliers. Kane's method \cite{Kane1985}
allows for formulation of equations of motion for constrained systems in which
no multipliers are present. The dynamics of these systems can be completely
described by ordinary differential equations. This significantly different
formulation requires a different approach to linearization than what the
existing options provide.

The goal of this paper is to concretely establish the first order
relationships (about an arbitrary equilibrium point) between a selection of
independent coordinates and speeds and the time-derivatives of all of the
coordinates and speeds (both independent and dependent). The formalism
presented here is generic enough to cover most examples of time-varying
constrained multibody systems with arbitrary external inputs and arbitrary
specified quantities.  This procedure was created for systems whose equations
of motion have been derived with Kane's Method. Although also applicable to
dynamic system equations formulated using other methods, it is restricted to
systems which can be completely described by a set of ODEs (i.e., DAEs needn't
be solved).

We first present our linearization procedure. Next we will offer some
discussion of the procedure, along with an example. The derivation of the
procedure is provided in Section \ref{sec:derivations}. Additionally, the electronic
supplementary material provides more examples (in more detail) and an
alternative ordering for the independent/dependent quantities.

@@

\section{Results}
\label{sec:results}
The linearized equations of motion are
\begin{align}
  \label{eq:state_space_constrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q}_i \\
        \delta \bm{u}_i
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n+m) \times s} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}
Table \ref{table:assumptions} provides the form in which the nonlinear
equations of motion and constraints must be arranged.  Equations
(\ref{eq:quant_to_compute}) provides the bulk of the terms in equation
(\ref{eq:state_space_constrained}); these terms are Jacobian matrices of the
differential equations. Equations (\ref{eq:C_0})-(\ref{eq:C_2}) provide the
terms $C_0$, $C_1$, and $C_2$ which embed the constraints into
(\ref{eq:state_space_constrained}). Finally, the permutation matrices in
equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the independent and
dependent speeds.

@@

\section{Discussion}
\label{sec:discussion}
Equation (\ref{eq:state_space_constrained}) doesn't fit into the standard linear
system framework because it is non-square.  To obtain a square system (useful
in stability analysis and control system design) we define
\begin{align}
  \label{eq:A_prime}
    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
  \label{eq:B_prime}
    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
\end{align}
where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
independent states by defining
\begin{align}
  \label{eq:P_prime}
    P^\prime &\triangleq \begin{bmatrix}
        P_{qi} & \bm{O}_{n \times (o - m)} \\
        \bm{O}_{o \times (n - l)} & P_{ui}
    \end{bmatrix} \\
  \label{eq:A}
    A &\triangleq P^{\prime T} A^\prime \\
  \label{eq:B}
    B &\triangleq P^{\prime T} B^\prime
\end{align}
where $P^\prime \in \bm{R}^{(o - m + n - l) \times (o - m + n - l)}$.  Defining
$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
linear systems analyses may applied.  It is worth noting that the rows
of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
in the output or measurement equations of a linear state space model, as in the
case when the controlling or measuring dependent state(s).

@@

%f_c:
%[r*cos(q1) + q5]
%f_v:
%[                r*u1*cos(q2) + u3]
%[-r*(u0*cos(q2) + u2*sin(q2)) + u4]
%[                r*u1*sin(q2) + u5]
%f_a:
%[                                -r*u1*sin(q2)*q2' + r*cos(q2)*u1' + u3']
%[-r*(-u0*sin(q2)*q2' + u2*cos(q2)*q2' + sin(q2)*u2' + cos(q2)*u0') + u4']
%[                                 r*u1*cos(q2)*q2' + r*sin(q2)*u1' + u5']
%f_0:
%[                                                                                      -sin(q2)*cos(q1)*q0' + cos(q2)*q1']
%[                                                                                                       sin(q1)*q0' + q2']
%[                                                                                       sin(q2)*q1' + cos(q1)*cos(q2)*q0']
%[(sin(q0)*cos(q2) + sin(q1)*sin(q2)*cos(q0))*q4' + (-sin(q0)*sin(q1)*sin(q2) + cos(q0)*cos(q2))*q3' - sin(q2)*cos(q1)*q5']
%[                                                                -sin(q0)*cos(q1)*q3' + sin(q1)*q5' + cos(q0)*cos(q1)*q4']
%[ (sin(q0)*sin(q2) - sin(q1)*cos(q0)*cos(q2))*q4' + (sin(q0)*sin(q1)*cos(q2) + sin(q2)*cos(q0))*q3' + cos(q1)*cos(q2)*q5']
%f_1:
%[-u0]
%[-u1]
%[-u2]
%[-u3]
%[-u4]
%[-u5]
%f_2:
%[                  -I*u0' - m*r*cos(q2)*u4']
%[-J*u1' + m*r*sin(q2)*u5' + m*r*cos(q2)*u3']
%[                  -I*u2' - m*r*sin(q2)*u4']
%f_3:
%[-I*u1*u2 + J*u1*u2 + g*m*r*sin(q1)*cos(q2) - m*r*(-u0*u5 + u2*u3)*cos(q2)]
%[                m*r*(u0*u4 - u1*u3)*sin(q2) + m*r*(u1*u5 - u2*u4)*cos(q2)]
%[ I*u0*u1 - J*u0*u1 + g*m*r*sin(q1)*sin(q2) - m*r*(-u0*u5 + u2*u3)*sin(q2)]

\begin{figure}
  \centering
  \input{rollingdisc.pgf}
  \caption{Rolling disc configuration}
  \label{fig:disc}
\end{figure}
\section{Derivation of linearization algorithm}
\label{sec:derivations}

This section outlines the derivation of the procedure for determining linear
equations of motion \ref{eq:state_space_constrained}. This equation can be used
for simulation, stability analysis, or it can be reformulated into the
traditional state space equations for a variety of tasks; these topics are
discussed in Section \ref{sec:discussion}.
